---
title: "HW5"
author: "Saatvik Anantha"
date: "2024-02-25"
output: pdf_document
---

<center>
EID: ska2242

GitHub Link: https://github.com/ska4/SDS315
</center>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(tidyverse)
library(knitr)
library(dplyr)
library(tibble)
library(boot)
```

\newpage

# Problem 1

Null Hypothesis: The rate of trades that are flagged at Iron Bank is being tested to see if it is the same as the baseline rate, which is 2.4%.

Test Statistic: Based on the null hypothesis, our test statistic will be the absolute difference between the observed and expected number of flagged trades, respectively.

```{r, echo=FALSE, include=TRUE}

test_stats <- numeric(100000)

for (i in 1:100000) {
  flags <- rbinom(1, 2021, 0.024)
  test_stats[i] <- abs(flags - 70)
}

num1 = 70 - 2021
num1 = num1 * 0.024

prob_1 <- hist(test_stats, col = "green", main = "Probability Distribution of Iron Bank Flagged Trade Test Statistic", xlab = "Test Statistic", ylab = "Frequency")
abline(v = abs(num1), col = "red")

p1 <- mean(test_stats >= abs(num1))

```

In the plot above, the observed test statistic hovers around the 47-48 range.

The P-value is about 2x10^-5, which is much less than 0.05.

Conclusion: Based on our findings, we can safely reject the null hypothesis and say that the rate of flagged trades is much different than the baseline rate for other traders, and I believe this seems very accurate due to the very low p-value.

\newpage

# Problem 2

Null Hypothesis: The rate of health code violations is being tested if it is the same as the average in the city, which is 3%, at Gourmet Bites.

Test Statistic: Based on the null hypothesis, our test statistic will be the absolute difference between the observed and expected number of health code violations, respectively.

```{r, echo=FALSE, include=TRUE}
test_stats <- numeric(100000)

for (i in 1:100000) {
  violations <- rbinom(1, 50, 0.03)
  test_stats[i] <- abs(violations - 8)
}

num2 = 8 - 50
num2 = num2 * 0.03

hist(test_stats, col = 'orange', main = "Probability Distribution of Gourmet Bites Health Inspection Test Statistic", xlab = "Test Statistic", ylab = "Frequency")
abline(v = abs(num2), col = "red")

p2 <- mean(test_stats >= abs(num2))
```

In the plot above, the observed test statistic hovers around the 1-1.3 range.

The P-value is about .9992, which is much greater than 0.05.

Conclusion: Based on our findings, we cannot reject the null hypothesis, and we can safely say that the rate of health code violations is pretty much the same as the citywide average, and I believe this seems very accurate due to the very high p-value.

\newpage

# Problem 3

```{r, echo=FALSE, include=TRUE}

## Part A

part_3a_chi_square <- function(sentence, letter_table) {
  
  letter_table$Probability <- letter_table$Probability/sum(letter_table$Probability)
  
  preprocess_sentence <- gsub("[^A-Za-z]", "", sentence)
  preprocess_sentence <- toupper(preprocess_sentence)
  
  letter_counts = table(factor(strsplit(preprocess_sentence, "")[[1]], levels = letter_table$Letter))
  
  letters = sum(letter_counts)
  expected_letter_counts = letters * letter_table$Probability
  
  chi_square <- sum((letter_counts - expected_letter_counts)^2 / expected_letter_counts)
  
  return(chi_square)
}

sentences_chi_squares <- numeric(0)
sentences <- readLines("brown_sentences.txt")
letter_frequencies <- read.csv("letter_frequencies.csv")

count = 1
for (i in sentences) {
  looped_sentence <- part_3a_chi_square(i, letter_frequencies)
  sentences_chi_squares[count] <- looped_sentence
  count = count + 1
}

## Part B


part_3b_chi_square <- function(sentence, letter_table, null_distribution) {
  
  letter_table$Probability <- letter_table$Probability/sum(letter_table$Probability)
  
  preprocess_sentence <- gsub("[^A-Za-z]", "", sentence)
  preprocess_sentence <- toupper(preprocess_sentence)
  
  letter_counts <- table(factor(strsplit(preprocess_sentence, "")[[1]], levels = letter_table$Letter))
  
  letters <- sum(letter_counts)
  
  expected_letter_counts <- letters * letter_table$Probability
  
  chi_square <- sum((letter_counts - expected_letter_counts)^2 / expected_letter_counts)
  p_value <- mean(null_distribution >= chi_square)
  
  return(list(chi_square = chi_square, p_value = p_value))
}

sentences <- readLines("brown_sentences.txt")
letter_frequencies <- read.csv("letter_frequencies.csv")
null_distribution <- numeric(0)
for (i in sentences) {
  looped_sentence <- part_3a_chi_square(i, letter_frequencies)
  null_distribution <- c(null_distribution, looped_sentence)
}

provided_sentences <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)

p_values <- numeric(length(provided_sentences))
chi_squares <- numeric(length(provided_sentences))

for (i in 1:length(provided_sentences)) {
  result <- part_3b_chi_square(provided_sentences[i], letter_frequencies, null_distribution)
  chi_squares[i] <- result$chi_square
  p_values[i] <- result$p_value
}

p_values_table <- data.frame(Sentence = provided_sentences, P_Value = round(p_values, 3))

kable(p_values_table, align = c("l", "r"), col.names = c("Sentence", "P-Value"), caption = "P-values for Given Ten Sentences")
```

I believe Sentence 6 is the sentence produced and watermarked by an LLM, as it has the lowest p-value of 0.009, and the sentence structure is longer and more complex comparatively.